2024 US Presidential Election Model

### 2024-10-19 paper/paper.qmd
saving code changes to paper - data section

library(tidyverse)
library(janitor)
library(lubridate)
library(modelsummary)
library(rstanarm)
library(readr)
library(dplyr)
```

@@

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

>>>>

library(tidyverse)
library(janitor)
library(lubridate)
library(modelsummary)
library(rstanarm)
library(readr)
library(dplyr)
library(ggplot2)
library(usmap)
```

@@

@fig-averagetable provides an overview of the average support for Trump and Harris across all states, along with the predicted winner for each state based on our model. 
```{r}
#| label: fig-averagetable
#| fig-cap: Summary Table of Polling Averages by State
#| echo: false

# Create winner column: 1 if Trump_pct is greater than Harris_pct, 0 otherwise
state_average <- state_average |>
  mutate(winner = ifelse(Trump_pct > Harris_pct, 1, 0))

average_table <- state_average |>
  select(state, Trump_pct, Harris_pct, winner) |>
  mutate(predicted_winner = ifelse(winner == 1, "Trump", "Harris"))

average_table
```

@fig-barplot illustrates the average polling support for Donald Trump and Kamala Harris in each state. The side-by-side comparison represents the percentage of support each candidate has received, based on aggregated poll data from pollsters with high-quality scores. Trump's support is shown in red, while Harris's support is depicted in red.

```{r}
#| label: fig-barplot
#| fig-cap: Average Poll Percentage by State
#| echo: false

# Reshape data to a long format for side-by-side bars
stateaverage_long <- state_average %>%
  pivot_longer(cols = c(Trump_pct, Harris_pct),
               names_to = "candidate",
               values_to = "support_pct")

ggplot(stateaverage_long, aes(x=reorder(state, -support_pct), y= support_pct, fill = candidate)) + 
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), color = "black") +
  labs(title = "Average Poll Percentages for Trump and Harris by State",
       x = "State", y = "Average Percentage") +
  scale_fill_manual(name = "Candidate", values = c("Trump_pct" = "red", "Harris_pct" = "blue")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

@fig-map visualizes the predicted winner of the 2024 US Presidential Election across all states, based on aggregated polling data. Each state is color-coded according to the candidate predicted to win - blue represents Donald Trump, and red represents Kamala Harris.

```{r}
#| label: fig-map
#| fig-cap: Predicted Election Winner by State
#| echo: false
#| warning: false
#| message: false

library(ggplot2)
library(usmap)

# Ensuring state names are in the correct format
state_average <- state_average %>%
  mutate(state = str_to_title(state))

# Creating binary variable winner column: 1 if Trump has more support, 0 if Harris has more support. 
state_average <- state_average %>%
  mutate(winner = ifelse(Trump_pct > Harris_pct, 1, 0))

# Creating predicted_winner column to label Trump and Harris
state_average <- state_average |>
  mutate(predicted_winner = ifelse(winner == 1, "Trump", "Harris"))

# Map visualization showing predicted winner by state
plot_usmap(data = state_average, values = "predicted_winner", regions = "state") + 
  scale_fill_manual(values = c("Trump" = "red", "Harris" = "blue"), name = "Predicted Winner") + 
  labs(title = "Predicted Winner of 2024 US Presidential Election by state") + 
  theme(legend.position = "right")
```
